Data-driven Reduction of Multiscale Stochastic Dynamical Systems

### 2014-07-11 reducible_SDE/paper/paper_draft.tex
edied multiscale paper

Multiscale SDE setting: fast/slow, want to recover slow variable(s) even when the fast variables are large/very noisy

\begin{equation}
\begin{aligned}
dx &= adt + dW_1\\
dy &= -\frac{y}{\epsilon} dt + \frac{1}{\sqrt{\epsilon}} dW_2
\end{aligned}
\end{equation}

So $y$ is a fast noise whose equilibrium measure is bounded and $\mathcal{O}(1)$

@@

Background on NIVs

\begin{equation}
dx_i = a(X) dt + dW_i
\end{equation}
and we will let $X = \begin{bmatrix} X^1 \\ X^2 \\ \vdots \\ X^n \end{bmatrix}$

We want to emphasize the link between multi scale SDEs and NIVs: NIVs rescales the variables so that the ``fast'' variables become small

saturation time of SDE has to be smaller than the timescales

Let $y = \frac{z}{\sqrt{\epsilon}}$. Then
\begin{equation}
\begin{aligned}
dx &= adt + dW_1\\
dz &= -\frac{z}{\epsilon} dt +  dW_2
\end{aligned}
\end{equation}

Therefore, now $z$ is a stochastic variable whose diffusion is unity, and whose equilibrium measure is $\mathcal{O}(\epsilon)$. 
%
Therefore, when $\epsilon \ll 1$, we have that $z \ll x$, and so we will recover $x$ only.

>>>>

Setting
\begin{itemize}
\item system of SDEs with separation of timescales
\item want to recover slow variable(s) even when the fast variables are have large variations
\end{itemize}

Consider equations of the form
\begin{equation} \label{eq:fast_slow_SDE}
\begin{aligned}
dx &= adt + dW^1\\
dy &= -\frac{y}{\epsilon} dt + \frac{1}{\sqrt{\epsilon}} dW^2
\end{aligned}
\end{equation}

So $y$ is a fast noise whose equilibrium measure is bounded and $\mathcal{O}(1)$.
%
We assume that the saturation time of fast scale  of the SDE is smaller than the other relevant timescales.

TODO: should we write the SDE in a more general form?

@@

Background on NIVs 	
\begin{itemize}

\item Want to recover ``intrinsic'' variables that are independent of the measurement.observation function
\item We define intrinsic as variables that have uncoupled noise with unit variance
\end{itemize}

\begin{equation}
dX^i = a(X) dt + dW^i
\end{equation}
and we will let $X = \begin{bmatrix} X^1 \\ X^2 \\ \vdots \\ X^n \end{bmatrix}$

We want to emphasize the link between multi scale SDEs and NIVs: NIVs rescales the variables so that the ``fast'' variables become small.
%
Consider the SDE in \eqref{eq:fast_slow_SDE}, and let $y = \frac{z}{\sqrt{\epsilon}}$. 
%
Then
\begin{equation}
\begin{aligned}
dx &= adt + dW_1\\
dz &= -\frac{z}{\epsilon} dt +  dW_2
\end{aligned}
\end{equation}
%
Therefore, $z$ is a stochastic variable whose diffusion is unity, and so $x$ and $z$ are the ``intrinsic variables'' for the underlying SDE.
%
However, the equilibrium measure of $z$ is $\mathcal{O}(\epsilon)$, and so when $\epsilon \ll 1$, $z \ll x$, and so we will only recover $x$ (the slow variable).
